Efficient Compilation of High Level Python Numerical Programs with Pythran

### 2014-09-03 doc/papers/cgo2014/paper.tex
rework rosen measure and description

As introduced in \S~\ref{sec:stackoverflow}, we use the numpy-benchmark test 
suite to evaluate the optimizations described in\S~\ref{sec:pythran}. 
Pythran is used (commit: 488b7ba) with and without SIMD optimization. 
The results are compared with Pypy \todo{add}(???), Parakeet (commit: 4e55810), 
Numba (commit: b8a79c5) and Python 2.7.

Benchmarking is done on two architectures with different backend compilers. The
\texttt{timeit} module is used to perform the measurments, taking the average
execution time over several runs.

The first one holds 8 AMD Magny Cours processors for a
total of 48 cores. Each core has access to 64\,KB of L1 cache,
512\,KB of L2 cache. Both L1 and L2 caches are private, while L3 cache
is shared between the 6 cores of a processor. This configuration
provides a total of 256\,GB (32\,GB per NUMA node) of main memory. It supports
up to SSE4. We refer to this configuration as \textbf{AMD48}.
On AMD48, the backend compilers are GCC (Debian 4.7.2-5) and the clang-omp
version from intel~\footnote{\url{http://clang-omp.github.io/}} to be able to
benefits from OpenMP improvements. This Clang version is based on Clang 3.4
release (LLVM:233b1e3f034, clang-omp:6f358bcd87).

\todo{je me demande si on devrait pas utiliser juste cette archi. 
Qu'est ce que l'autre apporte ? On sous utilise complètement les 48 coeurs...}
\todo{Mehdi: agree, sauf si on peut avoir UN exemple avec 48 threads qui scale 
"correctement"}

The second one holds one Intel(R) Xeon(R) CPU E5-2650 0 @ 2.00GHz for a total of
8 cores (multithreading is not used). Each node can access to 64\,KB of L1
cache, 256\,KB of L2 cache. Both L1 and L2 caches are private, while L3 cache
is shared between the 8 cores. This configuration provides a total of 64\,GB of 
main memory. It supports up to AVX. We refer to this configuration as
\textbf{Intel8}.
On Intel8, backend compilers is GCC 4.9 (20140528).

@@

% For performance comparison between different Pythran version
To compare the different optimization schemes, we run the numpy-benchmarks test
suite multiple times on the \textit{rosen} benchmark.
Figures~\ref{fig:rosenamd48} and \ref{fig:rosenintel8} illustrate the results.
The difference between the two architectures can be explain by AVX support on
the Intel8 architecture while AMD48 supports up to SSE4 only.
% Haaaaa, I don't understand why OMP is not use...
\todo{I have to redo mesures without forward substitution to show difference.}
% SG: may be the number of iteration is not enough? what array size did you use? (it should be specified in the text secion, BTW

@@

Figure~\ref{fig:arc_distance} is a typical example where we can make intensive
use of the techniques described in this paper. As we can see, both vectorization
and parallelization have an impact on execution time. We can notice that SIMD
contribution is not as good on AMD48 as it is on Intel8, because it does not
support AVX.
%FIXME: mehdi propose de n'utiliser qu'un backend pour centrer le propos sur Pythran/Numpy.
% ça allègera aussi les graphes...

>>>>

As introduced in \S~\ref{sec:stackoverflow}, we use the numpy-benchmark test 
suite to evaluate the optimizations described in\S~\ref{sec:pythran}. 
Pythran is used (commit: 488b7ba) with and without SIMD optimization. 
The results are compared with Pypy (commit: 632c832bd32f), Parakeet (commit: 4e55810), 
Numba (commit: b8a79c5) and Python 2.7.

numpy-benchmarks run 11 times benchmarks with an execution time around 1 seconde.
THe best time, average time and standard deviation is return from these 11 runs.

Benchmarking is done on two architectures with different backend compilers.

The first one holds 8 AMD Magny Cours processors for a
total of 48 cores. Each core has access to 64\,KB of L1 cache,
512\,KB of L2 cache. Both L1 and L2 caches are private, while L3 cache
is shared between the 6 cores of a processor. This configuration
provides a total of 256\,GB (32\,GB per NUMA node) of main memory. It supports
up to SSE4. We refer to this configuration as \textbf{AMD48}.
On AMD48, the backend compilers is the clang-omp
version from intel~\footnote{\url{http://clang-omp.github.io/}} to be able to
benefits from OpenMP improvements. This Clang version is based on Clang 3.4
release (LLVM:233b1e3f034, clang-omp:6f358bcd87).

The second one holds one Intel(R) Xeon(R) CPU E5-2650 0 @ 2.00GHz for a total of
8 cores (multithreading is not used). Each node can access to 64\,KB of L1
cache, 256\,KB of L2 cache. Both L1 and L2 caches are private, while L3 cache
is shared between the 8 cores. This configuration provides a total of 64\,GB of 
main memory. It supports up to AVX. We refer to this configuration as
\textbf{Intel8}.
On Intel8, backend compilers is GCC 4.9 (20140528) with libgomp.

@@

% For performance comparison between different Pythran version
To compare the different optimization schemes, we run the numpy-benchmarks test
suite multiple times on the \textit{rosen} benchmark. The algorithm is the one
on Listing~\ref{lst:rosen} and input is a raw array of 1,000,000 of double elements.
Figures~\ref{fig:rosenamd48} and \ref{fig:rosenintel8} illustrate the results.
The speedup difference between the two architectures can be explain by AVX support on
the Intel8 architecture while AMD48 supports up to SSE4 only.
We can also notice importance of forward substitution on this benchmarks as
we avoid extra loop, 2 temporary array assignment and avoid some load/store for
SIMD instructions.

@@

Figure~\ref{fig:arc_distance} is a typical example where we can make intensive
use of the techniques described in this paper. It measures performances on the
arc\_distance benchmark. It is a kernel with intensive points to points computation
and trigonometrics use. As we can see, both vectorization
and parallelization have an impact on execution time. We can notice that SIMD
contribution is not as good on AMD48 as it is on Intel8, because it does not
support AVX. This program also benefits from OpenMP. Input is 10000 elements 
double raw array.

%FIXME: mehdi propose de n'utiliser qu'un backend pour centrer le propos sur Pythran/Numpy.
% ça allègera aussi les graphes...
% On change juste le jeu de donnée plutot que de supprimer une archi
